Network Data Collection for Artifical Intelligence

### 2020-02-28 paper2/paper2.tex
Replace and remove cite

Before the network traffic can be analysed, it has to be recorded first. Therefore, this chapter will give an overview of the available third-party software which can be used to record all the incoming network traffic. The functional requirements for a tool are chiefly the pace of handling the network traffic. All of these tools are client-based tools. Means that the software needs a client to be set up and used. This means that either the network target for the traffic has to be a client or the e.g. switch has to clone a port with all the incoming and outgoing traffic and direct all the data to a client running one or more of these tools. In some cases, analysing of network traffic can take much longer than the data is send or received by the network which means, that all the data need to be buffered or stored. In this case a packet-to-disk packet recorder is needed. This device can capture large amount of network traffic with a data rate of up to 40 Gbit/s and write it into files.

@@

Every tools needs APIs\footnote{Application Programming Interface} to capture all the traffic sent through the network interface. Therefore, libraries were made for both Unix-like and Windows based operating systems. While \textit{libpcap} is used for Unix-like operating systems, a Windows system need the library called \textit{WinPcap}. With an pcap library software has the ability to tap all the network traffic from the selected interface. Based on the possibility to include the library in \textbf{C} and \textbf{C++}, \textit{pcap} can be used very easily into software. With the release of \textit{libpcap} 1.0.0, the zero-copy mechanism was implemented which allows the software to read the network traffic directly without copying the data from kernel space to user space.

@@

In this section we will itemise five well known and useful network recording tools. Some of the tools are console based tools, the other do have a GUI or both. All of these tools are client-based tools. Means that the software needs a Client to be set up and used. This means that either the network target for the traffic has to be a client or the e.g. switch has to clone a port with all the incoming and outgoing traffic and direct all the data to a client running one or more of these tools.

@@

Especially in the world of Unix based operating systems, \textit{tcpdump} is one of the best known tools for monitoring and sniffing network traffic. \textit{tcpdump} is exclusively command line based tool. This property makes \textit{tcpdump} a tool for advanced console line users only. But once you get familiar with the commands and options of \textit{tcpdump}, it will get a  a very efficient and powerful tool for many applications. \textit{tcpdump} uses the \textit{libpcap} network library of Unix based operating systems for capturing network traffic.

@@

With the wide set of options, \textit{tcpdump} is a pretty agile tool for dumping any analysing network traffic. A few of the options are listed below \cite{tcpdumporg}.

@@

If \textit{tcpdump} is compiled with cryptography, it is able to decrypt IPsec ESP packets in real time \cite{goyal2017}.
For the Windows enthusiasts a Windows derivative of \textit{tcpdump} is available. It is called \textit{WinDump}. \textit{WinDump} uses so called WinPcap library for Windows operating systems.  Which has the same APIs as libpcap for Unix-like systems. This means, network traffic captured, e.g. on a Ubuntu system can be analysed on a Windows 10 Client afterwards.

@@

\textit{Wireshark} is for good reason the best known cross-platform networking tool available. In 2006 \textit{Wireshark} was started as a fork of the software called \textit{Ethereal} by the Wireshark-Community due to trademark issues \cite{asrodiapatel2012}.  The performance of this tool is powerful and the set of functions is neatly arranged. By using GTK toolkit, \textit{Wireshark} has implemented a GUI \footnote{Graphical User Interface}\cite{suribatra2012}. The GUI is separated typically into three segmentations, traffic overview, details of selected packet and the raw data of each packet.

@@

A good asset of \textit{Wireshark} is its available set of filters for capturing and displaying network data. With the help of plugins, \textit{Wireshark} can even handle with new or dedicated protocols. Another big asset of \textit{Wireshark} is its capibility of capturing raw USB traffic from an USB-Port. By highlighting the traffic with different colours as shown in Fig.~\ref{wireshark} it's easier to scan for required traffic.

\subsection*{Netsniff-ng}

\textit{Netsniff} is a another Unix based tool for network analysis. \textit{Netsniff} was the first tool which had the ability to analyse the whole traffic with a zero-copy procedure, means that \textit{Netsniff} uses the same data as the kernel do. Based on this functional principle, \textit{Netsniff} is one of the best performing tools for network sniffing. \textit{Netsniff} is like \textit{tcpdump} exclusively an console based software. The set of options contains 38 capabilities on the latest version of \textit{Netsniff}: 0.6.4 (effective 26.02.2020).

@@

The fourth tool is called \textit{Capsa} from \textit{Colasoft} founded 2001 in United States. \textit{Capsa} is suitable for real-time network data capturing both LAN and Wireless Interfaces. A big advantage of \textit{Capsa} is the ability to evaluate the network traffic into graphical output like diagrams. This feature can help identifying perpetrator of inadvertent network traffic quickly by a human. The free version of \textit{Capsa} is heavily restricted regarding available functions.

@@

Yet another tool for network monitoring is called \textit{EtherApe}. \textit{EtherApe} is an Unix based software with a meaningful graphic interface and graphical data output of network traffic. With the help graphical illustration anomalies like a DDoS attack can be detected easily as shown in Fig.~\ref{etherape}. Every host is shown as a node in circle, every connection is shown as link between two nodes and every protocol has an own colour for the link. With higher network traffic between two hosts the link between nodes is shown larger.

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.37]{etherape.png}}
\caption{Graphic Representation of Network Traffic}
\label{etherape}
\end{figure}

@@

A common approach is to group packets into flows, and then calculate statistical features based on these flows. Generally, a flow is a group of packets with the same source and destination IP addresses and ports, and the same protocol \cite{brownlee1998}. However, a flow can also be terminated by a TCP FIN packet or a timeout \cite{Ids2018}. Statistical features of network flows include the following:
\begin{itemize}
	\item Flow duration \cite{Nb2015} \cite{Ids2018}
	\item Time between two flows or packets \cite{Ids2018}
	\item Time between SYN and SYN/ACK, and between SYN/ACK and ACK packets during TCP connection setup \cite{Nb2015}
	\item Active time before a flow becomes idle and vice versa \cite{Ids2018}
	\item Jitter \cite{Nb2015}
	\item Number of packets in total, with at least 1 byte of payload, or with certain protocol flags \cite{Ids2018} \cite{waizumi2007} \cite{celik2015}
	\item Number of fragmented packets \cite{waizumi2007}
	\item Number of retransmitted or dropped packets \cite{Nb2015}
	\item Size of packets, payload, headers or segments \cite{Ids2018} \cite{celik2015}
	\item Number of bytes/packets transferred per second, in bulk, or in the initial TCP window \cite{Ids2018} \cite{celik2015}
	\item Number of RTT samples \cite{celik2015}
	\item Time to live value \cite{Nb2015}
	\item TCP window advertisement value \cite{Nb2015}
	\item TCP base sequence number \cite{Nb2015}
	\item Ratio between download and upload \cite{Ids2018}
	\item Ratio between maximum and minimum packet size \cite{celik2015}
	\item Number of flows with the current flow's IP address or port number \cite{waizumi2007} \cite{Nb2015}
\end{itemize}
Most of these features can be derived by statistical calculations such as total, minimum, maximum, mean, median, standard deviation, variance, skewness and kurtosis \cite{Ids2018} \cite{celik2015, xu2015}. Also, the features can be extracted from bidirectional flows, or separately for forward and backward directions \cite{Ids2018}.

@@

\bibitem{b9} Nishant Kumar Singh, S. Thakur, H. Chaurasiya and H. Nagdev, Automated provisioning of application in IAAS cloud using Ansible configuration management, 1st International Conference on Next Generation Computing Technologies (NGCT), Dehradun, 2015, pp. 81-85.

@@

@@

\bibitem{brownlee1998} Brownlee, N. (1998). Network management and realtime traffic flowmeasurement. J. Netw. Syst. Manage., vol. 6, pp. 223–228, June 1998.

>>>>

Before the network traffic can be analysed, it has to be recorded first. Therefore, this chapter will give an overview of the available third-party software which can be used to record all the incoming network traffic. The functional requirements for a tool are chiefly the pace of handling the network traffic. All of these tools are client-based tools. This means that either the network target for the traffic has to be a client or the e.g. switch has to clone a port with all the incoming and outgoing traffic and direct all the data to a client running one or more of these tools. In some cases, analysing of network traffic can take much longer than the data is sent or received by the network which means, that all the data need to be buffered or stored. In this case a packet-to-disk recorder is needed. This device can capture large amount of network traffic with a data rate of up to 40 Gbit/s and write it into files.

@@

Every tools needs APIs\footnote{Application Programming Interface} to capture all the traffic sent through the network interface. Therefore, libraries were made for both Unix-like and Windows based operating systems. While \textit{libpcap} is used for Unix-like operating systems, a Windows system needs the library called \textit{WinPcap}. In 2013 the last release (4.1.3) of \textit{WinPcap} was published. Since then \textit{Npcap} acts as a substitute\cite{winpcaporg}. With a pcap library, software has the ability to tap all the network traffic from the selected interface. Based on the possibility to include the library in \textbf{C} and \textbf{C++}, \textit{pcap} can be used very easily in software. With the release of \textit{libpcap} 1.0.0, the zero-copy mechanism was implemented which allows the software to read the network traffic directly without copying the data from kernel space to user space.

@@

In this section we will itemise five well known and useful network recording tools. Some of the tools are console based tools, others have a GUI, or both.

@@

Especially in the world of Unix based operating systems, \textit{tcpdump} is one of the best known tools for monitoring and sniffing network traffic. \textit{tcpdump} is an exclusively command line based tool. This property makes \textit{tcpdump} a tool for advanced command line users only. But once you get familiar with the commands and options of \textit{tcpdump}, it will get a  a very efficient and powerful tool for many applications. \textit{tcpdump} uses the \textit{libpcap} network library of Unix based operating systems for capturing network traffic.

@@

With the wide set of options, \textit{tcpdump} is a pretty agile tool for dumping and analysing network traffic. A few of the options are listed below \cite{tcpdumporg}.

@@

If \textit{tcpdump} is compiled with cryptography, it is able to decrypt IPsec ESP packets in real time \cite{goyal2017}.
For the Windows enthusiasts a Windows derivative of \textit{tcpdump} is available. It is called \textit{WinDump}. \textit{WinDump} uses the \textit{WinPcap} library, which provides the same API for Windows operating systems as \textit{libpcap} for Unix-like systems. This means, network traffic captured, e.g. on a Ubuntu system can be analysed on a Windows 10 Client afterwards.

@@

\textit{Wireshark} is for good reason the best known cross-platform networking tool available. In 2006 \textit{Wireshark} was started as a fork of the software called \textit{Ethereal} by the Wireshark-Community due to trademark issues \cite{asrodiapatel2012}.  The performance of this tool is powerful and the set of functions is neatly arranged. By using the GTK toolkit, \textit{Wireshark} has implemented a GUI \footnote{Graphical User Interface}\cite{suribatra2012}. The GUI is separated typically into three segments: the traffic overview, and the details and raw data of the selected packet.

@@

A good asset of \textit{Wireshark} is its available set of filters for capturing and displaying network data. With the help of plugins, \textit{Wireshark} can even handle with new or dedicated protocols. Another big asset of \textit{Wireshark} is its capability of capturing raw USB traffic from an USB port. By highlighting the traffic with different colours as shown in Fig.~\ref{wireshark} it's easier to scan for required traffic.

\subsection*{netsniff-ng}

\textit{netsniff-ng} is a another Unix based tool for network analysis. \textit{netsniff-ng} was the first tool which had the ability to analyse the whole traffic with a zero-copy procedure, which means that \textit{netsniff-ng} uses the same data as the kernel does. Based on this functional principle, \textit{netsniff-ng} is one of the best performing tools for network sniffing. \textit{netsniff-ng} is like \textit{tcpdump} exclusively an console based software. The set of options contains 38 capabilities on the latest version 0.6.4 of \textit{netsniff-ng}.

@@

The fourth tool is called \textit{Capsa} from \textit{Colasoft} founded 2001 in United States. \textit{Capsa} is suitable for real-time network data capturing both LAN and Wireless Interfaces. A big advantage of \textit{Capsa} is the ability to evaluate the network traffic into graphical output like diagrams. This feature can help identifying the source of undesirable network traffic quickly by a human. The free version of \textit{Capsa} is heavily restricted regarding available functions.

@@

Yet another tool for network monitoring is called \textit{EtherApe}. \textit{EtherApe} is an Unix based software with a meaningful graphic interface and graphical data output of network traffic. With the help of graphical illustration anomalies like a DDoS attack can be detected easily as shown in Fig.~\ref{etherape}. Every host is shown as a node in circle, every connection is shown as link between two nodes and every protocol has an own colour for the link. With higher network traffic between two hosts the link between nodes is shown larger.

\begin{figure}[htbp]
\centerline{\includegraphics[scale=0.37]{etherape.png}}
\caption{Graphic Representation of Network Traffic \cite{etherapewiki}}
\label{etherape}
\end{figure}

@@

A common approach is to group packets into flows, and then calculate statistical features based on these flows. Generally, a flow is a group of packets with the same source and destination IP addresses and ports, and the same protocol. However, a flow can also be terminated by a TCP FIN packet or a timeout \cite{Ids2018}. Statistical features of network flows include the following:
\begin{itemize}
	\item Flow duration \cite{Nb2015} \cite{Ids2018}
	\item Time between two flows or packets \cite{Ids2018}
	\item Time between SYN and SYN/ACK, and between SYN/ACK and ACK packets during TCP connection setup \cite{Nb2015}
	\item Active time before a flow becomes idle and vice versa \cite{Ids2018}
	\item Jitter \cite{Nb2015}
	\item Number of packets in total, with at least 1 byte of payload, or with certain protocol flags \cite{Ids2018} \cite{waizumi2007} \cite{celik2015}
	\item Number of fragmented packets \cite{waizumi2007}
	\item Number of retransmitted or dropped packets \cite{Nb2015}
	\item Size of packets, payload, headers or segments \cite{Ids2018} \cite{celik2015}
	\item Number of bytes/packets transferred per second, in bulk, or in the initial TCP window \cite{Ids2018} \cite{celik2015}
	\item Number of RTT samples \cite{celik2015}
	\item Time to live value \cite{Nb2015}
	\item TCP window advertisement value \cite{Nb2015}
	\item TCP base sequence number \cite{Nb2015}
	\item Ratio between download and upload \cite{Ids2018}
	\item Ratio between maximum and minimum packet size \cite{celik2015}
	\item Number of flows with the current flow's IP address or port number \cite{waizumi2007} \cite{Nb2015}
\end{itemize}
Most of these features can be derived by statistical calculations such as total, minimum, maximum, mean, median, standard deviation, variance, skewness and kurtosis \cite{Ids2018} \cite{celik2015, xu2015}. Also, the features can be extracted from bidirectional flows, or separately for forward and backward directions \cite{Ids2018}.

@@

\bibitem{b9} Sandobalín, J. and Insfran, E. and Abrahão, S. (2020). On the Effectiveness of Tools to Support Infrastructure as Code: Model-Driven versus Code-Centric. IEEE Access. 10.1109/ACCESS.2020.2966597.

@@

\bibitem{winpcaporg} WinPcap: News, Riverbed Technology (2010), https://www.winpcap.org/misc/copyright.htm [Accessed 26.02.2020]
\bibitem{etherapewiki} https://upload.wikimedia.org/wikipedia/commons/a/ae/Etherape-gui.png [Accessed 25.02.2020]

@@
